---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# AdjKM.CIF

<!-- badges: start -->
<!-- badges: end -->

The goal of AdjKM.CIF is to create the covariate-adjusted Kaplan-Meier and cumulative incidence functions.

You could check the more specific introduction of the package at <https://lesly1031.github.io/AdjKM.CIF/>

## Installation

You can install the development version of AdjKM.CIF from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Lesly1031/AdjKM.CIF",dependencies = TRUE)
```

## Adjusted KM examples

```{r,warning=FALSE,message=FALSE}
library(AdjKM.CIF)
library(tidyverse)
library(DT)
library(data.table)
library(KMsurv)
```

### Example data

```{r}
data(bmt)
bmt$arm <- bmt$group
bmt$arm = factor(as.character(bmt$arm), levels = c("2", "1", "3"))
bmt$z3 = as.character(bmt$z3)
bmt$t2 = bmt$t2 * 12/365.25
```

### Coxph model

```{r}
result1 = adjusted_KM(data = bmt, time = "t2", status = "d3", group = "arm", covlist = c("z1",
    "z3"), stratified_cox = "No", reference_group = NULL)
```

```{r}
adjKM_plot(result1, data = bmt)
```

### Bootstrap results for Cox model

```{r}
result1_1 = boot_ci_adj_km(boot_n = 100, ci_cut = c(0.025, 0.975), data = bmt, time = "t2",
    status = "d3", group = "arm", covlist = c("z1", "z3"), stratified_cox = "No",
    reference_group = NULL)
```

```{r}
adjKM_CI_plot(result1_1, bmt)
```

## Adjusted CIF Examples

### Example data

```{r}
data(bmt)
bmt$arm <- bmt$group
bmt$arm = factor(as.character(bmt$arm), levels = c("2", "1", "3"))
bmt$z3 = as.character(bmt$z3)

bmt$CenCI <- 0

for (ii in 1:137) {
    if (bmt$d3[ii] == 0) {
        bmt$CenCI[ii] = 0
    } else {
        if (bmt$d2[ii] == 1) {
            bmt$CenCI[ii] <- 1
        } else {
            bmt$CenCI[ii] <- 2
        }
    }
}
bmt$t2 = bmt$t2 * 12/365.25
```


### FG model for event = relapse

```{r,warning=FALSE,message=FALSE}
result1 = adjusted_CIF(data = bmt, time = "t2", status = "CenCI", group = "arm",
    covlist = c("z1", "z3"), event_code = 1, stratified = "No", reference_group = NULL)
table_res1 = spread(result1, class, prob)
head(table_res1)
```

> Figure

```{r}
adjCIF_plot(result1, data = bmt)
```

### Bootstrap results for FG model with event = relapse

```{r,warning=FALSE,message=FALSE}
result1_boot = boot_ci_adj_cif(boot_n = 100, ci_cut = c(0.025, 0.975), data = bmt, time = "t2",
    status = "CenCI", group = "arm", covlist = c("z1", "z3"), event_code = 1, "No",
    NULL)
```

> Figure 

```{r}
adjCIF_CI_plot(result1_boot, bmt)
```



