---
title: "package_tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{package_tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup,warning=FALSE,message=FALSE}
library(AdjKM.CIF)
library(tidyverse)
library(DT)
library(data.table)
library(KMsurv)
```

# Covariate-Adjusted KM curves and probability

## Example data

```{r}
data(bmt)
bmt$arm <- bmt$group
bmt$arm = factor(as.character(bmt$arm),levels=c("2","1","3"))
#bmt$z10 = as.character(bmt$z10)
bmt$z3 = as.character(bmt$z3)
bmt$t2 = bmt$t2*12/365.25
```

## Adjusted KM

### Coxph model

```{r}
result1 = adjusted_KM (data = bmt,time = "t2", status = "d3",group = "arm",
                       covlist = c("z1","z3"),stratified_cox = "No", reference_group = NULL)
table_res1 = spread(result1, class, prob)
head(table_res1)
```


```{r}
adjKM_plot(result1,data=bmt)
```

> Alternative: use ggplot to finalize figure

Users can also use ggplot to customize figures as needed (e.g.change x ticks to every 6 months, change lab names)

```{r}
ggplot(result1,aes(x=time,y = prob, group =class))+
      geom_step(aes(linetype=class,color = class),size=1.5)+
      theme_classic()+
      ylim(c(0,1))+
      scale_x_continuous(name="Time (Months)",breaks = seq(0,84,6))+
      ylab("Probability")
```

> save results

```{r,eval = FALSE}
write.csv(table_res1,"cox_KM_res.csv")
```

### Stratified Cox model:  Gail & Byar's approach

```{r}
result2 = adjusted_KM (data = bmt,time = "t2", status = "d3",group = "arm",
                       covlist = c("z1","z3"),stratified_cox = "Yes", reference_group = "G&B")

table_res2 = spread(result2, class, prob)
head(table_res2)
```

```{r}
adjKM_plot(result2,data=bmt)
```

> Alternative: use ggplot to finalize figure


```{r}
ggplot(result2,aes(x=time,y = prob, group =class))+
      geom_step(aes(linetype=class,color = class),size=1.5)+
      theme_classic()+
      ylim(c(0,1))+
      scale_x_continuous(name="Time (Months)",breaks = seq(0,84,6))+
      ylab("Probability")

```

> save results

```{r,eval = FALSE}
write.csv(table_res2,"GB_KM_res.csv")
```

### Stratified Cox model: Storer's approach

```{r}
result3 = adjusted_KM (data = bmt,time = "t2", status = "d3",group = "arm", 
                       covlist = c("z1","z3"),stratified_cox = "Yes", reference_group = "arm:2")

table_res3 = spread(result3, class, prob)
head(table_res3)
```

```{r}
adjKM_plot(result3,data=bmt)
```

> Alternative: use ggplot to finalize figure


```{r}
ggplot(result3,aes(x=time,y = prob, group =class))+
      geom_step(aes(linetype=class,color = class),size=1.5)+
      theme_classic()+
      ylim(c(0,1))+
      scale_x_continuous(name="Time (Months)",breaks = seq(0,84,6))+
      ylab("Probability")

```

> save results

```{r,eval = FALSE}
write.csv(table_res3,"storer_KM_res.csv")
```

## Bootstrap CI

### Cox PH model

```{r}
result1_1 = boot_ci_adj_km(boot_n=100,ci_cut=c(0.025,0.975),data = bmt, time = "t2", status = "d3",group = "arm",
                         covlist = c("z1","z3"), stratified_cox = "No", reference_group = NULL)
```

> Figure

```{r}
adjKM_CI_plot(result1_1,bmt)
```

```{r}
result1 = do.call(rbind,result1_1)

ggplot(data.frame(result1),aes(x=time,y = t_sub_mean, group =class ))+
  geom_step(aes(linetype=class,color = class),size=1.2)+
  geom_ribbon(aes(ymin=lower,ymax=upper,fill=class),alpha=0.3)+
  ylim(c(0,1))+
  theme_classic()+
  scale_x_continuous(name = "Time (Months)",
    breaks = seq(0, 84, 6)) +
    ylab("Probability")
```

> save bootstrap results

```{r,eval=FALSE}
res_km = do.call(rbind,result1_1)
write.csv(res_km,"boots_km_cox.csv")
```

### Stratified Cox: G&B for event = relapse {.tabset}

```{r,include=TRUE}
result1_2 =  boot_ci_adj_km(boot_n=100,ci_cut=c(0.025,0.975),data = bmt, time = "t2", status = "d3",group = "arm",
                         covlist = c("z1","z3"), stratified_cox = "Yes", reference_group = "G&B")

```
    
```{r}
adjKM_CI_plot(result1_2,bmt)
```

```{r}
result2 = do.call(rbind,result1_2)

ggplot(data.frame(result2),aes(x=time,y = t_sub_mean, group =class ))+
  geom_step(aes(linetype=class,color = class),size=1.2)+
  geom_ribbon(aes(ymin=lower,ymax=upper,fill=class),alpha=0.3)+
  ylim(c(0,1))+
  theme_classic()+
  scale_x_continuous(name = "Time (Months)",
    breaks = seq(0, 84, 6)) +
    ylab("Probability")
    
```

> save bootstrap results

```{r,eval=FALSE}
res_km = do.call(rbind,result1_2)
write.csv(res_km,"boots_km_cox.csv")
```

### Stratified Cox: Storer for event = relapse {.tabset}

```{r,include=TRUE}
result1_3 =  boot_ci_adj_km(boot_n=100,ci_cut=c(0.025,0.975),data = bmt, time = "t2", status = "d3",group = "arm",
                         covlist = c("z1","z3"), stratified_cox = "Yes", reference_group = "arm:2")

```

> Figure


```{r}
adjKM_CI_plot(result1_3,bmt)
```

```{r}
result3 = do.call(rbind,result1_3)

ggplot(data.frame(result3),aes(x=time,y = t_sub_mean, group =class ))+
  geom_step(aes(linetype=class,color = class),size=1.2)+
  geom_ribbon(aes(ymin=lower,ymax=upper,fill=class),alpha=0.3)+
  ylim(c(0,1))+
  theme_classic()+
  scale_x_continuous(name = "Time (Months)",
    breaks = seq(0, 84, 6)) +
    ylab("Probability")
    
```


> save bootstrap results

```{r,eval=FALSE}
res_km = do.call(rbind,result1_3)
write.csv(res_km,"boots_km_cox.csv")
```

# Covariate-Adjusted CIF curves and probability

## Example data

```{r}
data(bmt)
bmt$arm <- bmt$group
bmt$arm = factor(as.character(bmt$arm),levels=c("2","1","3"))
bmt$z3 = as.character(bmt$z3)

bmt$CenCI <- 0
for(ii in 1:137) {
  if (bmt$d3[ii] == 0){
    bmt$CenCI[ii] <- 0
    }else {
    if (bmt$d2[ii] == 1){
      bmt$CenCI[ii] <- 1
  }else{
      bmt$CenCI[ii] <- 2}}
}
bmt$t2 = bmt$t2*12/365.25
```

## Adjusted CIF

## FG model for event = relapse

```{r}
result1 =adjusted_CIF(data = bmt,time = "t2", status = "CenCI",group = "arm", covlist = c("z1","z3"),event_code=1,stratified = "No", reference_group = NULL)
table_res1 = spread(result1, class, prob)
head(table_res1)
```

```{r}
adjCIF_plot(result1,data=bmt)
```

> Use ggplot to finalize figure

Users can also use ggplot to customize figures as needed (e.g.change x ticks to every 6 months, change lab names)

```{r}
ggplot(result1,aes(x=time,y = prob, group =class))+
      geom_step(aes(linetype=class,color = class),size=1.5)+
      theme_classic()+
      ylim(c(0,1))+
      scale_x_continuous(name="Time (Months)",breaks = seq(0,84,6))+
      ylab("Probability")

```

> save results

```{r,eval = FALSE}
write.csv(table_res1,"cox_CIF_res.csv")
```

### Stratified FG: G&B approach for event = relapse

```{r}
result2 = adjusted_CIF(data = bmt,time = "t2", status = "CenCI",group = "arm",covlist = c("z1","z3"),event_code=1,stratified = "Yes", reference_group = "G&B")
table_res2 = spread(result2, class, prob)
head(table_res2)
```

```{r}
adjCIF_plot(result2,data=bmt)
```

> save results

```{r,eval = FALSE}
write.csv(table_res2,"GB_CIF_res.csv")
```

> Use ggplot to finalize figure

```{r}
ggplot(result2,aes(x=time,y = prob, group =class))+
      geom_step(aes(linetype=class,color = class),size=1.5)+
      theme_classic()+
      ylim(c(0,1))+
      scale_x_continuous(name="Time (Months)",breaks = seq(0,84,6))+
      ylab("Probability")
```

### Stratified FG: Storer's approach for event = relapse

```{r}
result3 = adjusted_CIF (data = bmt,time = "t2", status = "CenCI",group = "arm",covlist = c("z1","z3"),event_code=1,stratified = "Yes", reference_group = "arm:2")

table_res3 = spread(result3, class, prob)
datatable(table_res3)%>%
  formatRound(columns=c("1","2","3"),digits = 3)
```

```{r}
adjCIF_plot(result3,data=bmt)
```


> Alternative: use ggplot to finalize figure


```{r}
ggplot(result3,aes(x=time,y = prob, group =class))+
      geom_step(aes(linetype=class,color = class),size=1.5)+
      theme_classic()+
      ylim(c(0,1))+
      scale_x_continuous(name="Time (Months)",breaks = seq(0,84,6))+
      ylab("Probability")
```


> save results

```{r,eval = FALSE}
write.csv(table_res3,"storer_CIF_res.csv")
```

## Bootstrap CI

### FG model for event = relapse

```{r,include=TRUE}
result1_1 = boot_ci_adj_cif(boot_n=100,ci_cut=c(0.025,0.975),data = bmt,time = "t2", status = "CenCI",group = "arm",covlist = c("z1","z3"),event_code=1,"No",NULL)
```

> Figure


```{r}
adjCIF_CI_plot(result1_1,bmt)
```

```{r}
result1 = do.call(rbind,result1_1)

ggplot(data.frame(result1),aes(x=time,y = t_sub_mean, group =class ))+
  geom_step(aes(linetype=class,color = class),size=1.2)+
  geom_ribbon(aes(ymin=lower,ymax=upper,fill=class),alpha=0.3)+
  ylim(c(0,1))+
  theme_classic()+
  scale_x_continuous(name = "Time (Months)",
    breaks = seq(0, 84, 6)) +
    ylab("Probability")
    
```


> save bootstrap results

```{r,eval=FALSE}
res_boot = do.call(rbind,result1_1)
write.csv(res_boot,"boots_cif_cox.csv")
```

### Stratified FG: G&B for event = relapse 

```{r,include=TRUE}
result1_2 = boot_ci_adj_cif(boot_n=100,ci_cut=c(0.025,0.975),data = bmt,time = "t2", status = "CenCI",group = "arm",covlist = c("z1","z3"),event_code=1,"Yes","G&B")

```

> Figure

```{r}
adjCIF_CI_plot(result1_2,bmt)
```

```{r}
result2 = do.call(rbind,result1_2)

ggplot(data.frame(result2),aes(x=time,y = t_sub_mean, group =class ))+
  geom_step(aes(linetype=class,color = class),size=1.2)+
  geom_ribbon(aes(ymin=lower,ymax=upper,fill=class),alpha=0.3)+
  ylim(c(0,1))+
  theme_classic()+
  scale_x_continuous(name = "Time (Months)",
    breaks = seq(0, 84, 6)) +
    ylab("Probability")
    
```

> save bootstrap results

```{r,eval=FALSE}
res_boot = do.call(rbind,result1_2)
write.csv(res_boot,"boots_cif_cox.csv")
```

### Stratified FG: Storer for event = relapse

```{r,include=TRUE}
# unstratified Fine-Gray regression model 
result1_3 = boot_ci_adj_cif(boot_n=100,ci_cut=c(0.025,0.975),data = bmt,time = "t2", status = "CenCI",group = "arm",covlist = c("z1","z3"),event_code=1,"Yes","arm:2")
```

> Figure

```{r}
adjCIF_CI_plot(result1_3,bmt)
```


```{r}
result3 = do.call(rbind,result1_3)

ggplot(data.frame(result3),aes(x=time,y = t_sub_mean, group =class ))+
  geom_step(aes(linetype=class,color = class),size=1.2)+
  geom_ribbon(aes(ymin=lower,ymax=upper,fill=class),alpha=0.3)+
  ylim(c(0,1))+
  theme_classic()+
  scale_x_continuous(name = "Time (Months)",
    breaks = seq(0, 84, 6)) +
    ylab("Probability")
    
```

> save bootstrap results

```{r,eval=FALSE}
res_boot = do.call(rbind,result1_3)
write.csv(res_boot,"boots_cif_cox.csv")
```
